Classless Java: Tuning Java Interfaces

### 2015-11-26 paper/sections/RelatedWork.tex
related work before 8.3

\paragraph{C++ Approach.}
C++ tries to provide a general solution to multiple inheritance. Virtual
inheritance in C++ provides another solution to multiple inheritance (especially
the diamond problem by keeping only one copy of the base
class)~\cite{ellis1990annotated}, however suffers from object initialization
problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
constructor calls to virtual superclasses, which would potentially cause serious
semantic errors. \bruno{What happens in our approach for the same case?}

@@

\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse (traits) and object factories
(classes). In this model, traits as units of reusable code, contain
only methods as reusable entities. Thus, no state and state
initialization are considered. Classes act as object factories,
requiring functionalities from multiple traits. Java 8 interfaces with
default methods are closely related to traits: concrete method
implementation are allowed (via the \textbf{default} keyword) inside
interfaces. The introduction of default methods opens the gate for
various flavours of multiple inheritance in Java, using interfaces.
Traits offer a trait algebra with operations like sum, alias and exclusion, provided
for explicit conflict resolution.  
Former work by Bono et al.~\cite{bono14}. provides details on
mimicking the trait algebra through Java 8 interfaces. 
We briefly recall the main points of their encoding;
however we propose a different representation of \textbf{exclusion}.
The author of ~\cite{bono14} agree that our revised version is
cleaner, typesafe and more direct.\bruno{do we want do say this?}

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\bruno{who?} see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

@@

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance. However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is a big limitation in class based object oriented
programming, and is exasperated by the interface-based programming we propose: no
type is ever going to be exact since classes are not explicitly used. A recent
article~\cite{Saito2013933} lights up this topic, proposing two
new features: exact statements and non-inheritable methods. Both are
related to our work: any method generated inside of the \Q@of@ method is indeed
non-inheritable, since there is no class name to extend from, and exact
statements (a form of wild-card capture on the exact run-time type) could
capture the ``exact type'' of an object even in a class-less environment.
%Interestingly, our use of covariant type refinement is enough to emulate

>>>>

\paragraph{C++ Approach.}
C++ tries to provide a general solution to multiple inheritance. Virtual
inheritance in C++ provides another solution to multiple inheritance (especially
the diamond problem by keeping only one copy of the base
class)~\cite{ellis1990annotated}, however suffers from object initialization
problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
constructor calls to virtual superclasses, which would potentially cause serious
semantic errors. \bruno{What happens in our approach for the same
  case?}\yanlin{explained below please double-check.} In our
approach, \mixin annotation processor has the full control of object
initialization mechanism, and this mechanism is clean and transparent to
users. Moreover, if not satisfied with the default generated \texttt{of} method,
customized factory methods are also allowed to be defined.

@@

\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse (traits) and object factories
(classes). In this model, traits as units of reusable code, contain
only methods as reusable entities. Thus, no state and state
initialization are considered. Classes act as object factories,
requiring functionalities from multiple traits. Java 8 interfaces with
default methods are closely related to traits: concrete method
implementation are allowed (via the \textbf{default} keyword) inside
interfaces. The introduction of default methods opens the gate for
various flavours of multiple inheritance in Java, using interfaces.
Traits offer a trait algebra with operations like sum, alias and exclusion, provided
for explicit conflict resolution.  
Former work by Bono et al.~\cite{bono14}. provides details on
mimicking the trait algebra through Java 8 interfaces. 
We briefly recall the main points of their encoding;
however we propose a different representation of \textbf{exclusion}.
The author of ~\cite{bono14} agree that our revised version is
cleaner, typesafe and more direct.\bruno{do we want do say this?}\yanlin{need to
  ask Marco.}

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\bruno{who?}\yanlin{need to ask Marco} see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

@@

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance. However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is a big limitation in class based object oriented
programming, and is exasperated by the interface-based programming we propose: no
type is ever going to be exact since classes are not explicitly used. A recent
article~\cite{Saito2013933} lights up this topic, proposing two
new features: exact statements and nonheritable methods. Both are
related to our work: 1) any method generated inside of the \Q@of@ method is indeed
non-inheritable since there is no class name to extend from; 2)exact
statements (a form of wild-card capture on the exact run-time type) could
capture the ``exact type'' of an object even in a class-less environment.
%Interestingly, our use of covariant type refinement is enough to emulate
